---
title: "Shetland CV LTC"
author: "James McMahon and Sue Wakely"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  html_document:
    css: phs_style.css
    toc: true
    toc_float: true
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Copy in the PHS style CSS file if required
if (!file.exists(file.path(getwd(), "phs_style.css"))) {
  phstemplates::add_stylecss(getwd(), auto_open = FALSE, shinycss = FALSE)
}
```

```{r load_packages, include = FALSE}
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(scales) # Scale Functions for Visualization
library(phsstyles) # Standard Graphic Styles for use in Public Health Scotland
```

# CV disease LTC - Read Code analysis

*Note that this project is a work in progress and the outputs are **not for onward sharing or distribution***

## Prevalence

```{r get_prev_data, include=FALSE}
source("prevalence.R")
```

The CV prevalence for the most recent three years, this is the monthly prevalence as a rate per 1,000 list size.

```{r cv_prev_3_year_list}
# List size recent 3 year
monthly_summary |>
  filter(census_date >= as.Date("2021-11-01")) |>
  ggplot(aes(x = census_date, y = list_prev * 1000)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(
    breaks = breaks_pretty(),
    labels = label_number(),
    limits = c(0, NA)
  ) +
  scale_x_date(
    breaks = breaks_pretty(10),
    labels = label_date("%b %y")
  ) +
  theme_minimal() +
  labs(
    title = "CV LTC prevalence in the most recent 3 years",
    x = "Census date",
    y = "Prevalence per 1,000 list size"
  )
```

We can look back at the 'prevalence' across all the time, but historic data completeness is insufficient to draw any conclusions.

```{r cv_prev_full_pops}
# Full using populations
monthly_summary |>
  ggplot(aes(x = census_date, y = pop_prev * 1000)) +
  geom_line() +
  scale_y_continuous(
    breaks = breaks_pretty(),
    labels = label_number(),
    limits = c(0, NA)
  ) +
  scale_x_date(
    breaks = breaks_pretty(10),
    labels = label_date("%b %y")
  ) +
  theme_minimal() +
  labs(
    title = "CV LTC prevalence over time",
    subtitle = "Older data is incomplete",
    x = "Census date",
    y = "Prevalence per 1,000 population"
  )
```

## LTC Invites

Within the last three years, at each month (census date), what percentage of the patients diagnosed with CV had received an invite within the preceding 15 months.

```{r ltc_invite}
# List size recent 3 year
monthly_summary |>
  filter(census_date >= as.Date("2021-11-01")) |>
  ggplot(aes(x = census_date, y = ltc_invite_prop)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(
    breaks = breaks_pretty(),
    labels = label_percent(),
    limits = c(0, NA)
  ) +
  scale_x_date(
    breaks = breaks_pretty(10),
    labels = label_date("%b %y")
  ) +
  theme_minimal() +
  labs(
    title = "LTC Invite Percent",
    subtitle = str_wrap("The percentage of patients with a CV diagnosis who had an Invite in the 15 months preceding the census date", width = 50),
    x = "Census date",
    y = "Percent of CV diagnosed patients"
  )
```

## LTC Attendances

Within the last three years, at each month (census date), what percentage of the patients diagnosed with CV had attended within the preceding 15 months.

```{r ltc_attend}
# List size recent 3 year
monthly_summary |>
  filter(census_date >= as.Date("2021-11-01")) |>
  ggplot(aes(x = census_date, y = ltc_attend_prop)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(
    breaks = breaks_pretty(),
    labels = label_percent(),
    limits = c(0, NA)
  ) +
  scale_x_date(
    breaks = breaks_pretty(10),
    labels = label_date("%b %y")
  ) +
  theme_minimal() +
  labs(
    title = "LTC Attendance Percent",
    subtitle = str_wrap("The percentage of patients with a CV diagnosis who had an Attendance in the 15 months preceding the census date", width = 50),
    x = "Census date",
    y = "Percent of CV diagnosed patients"
  )
```

